\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-11-01 paper.tex
Updates from Overleaf

@@

To address the feasibility challenge and performance trade-off of the ADMM problem discussed in \Cref{sec:reachability-discussion}, we propose to form \emph{sub-teams} on each route, and setup additional co-observations both within the sub-team and across different sub-teams (as shown in \Cref{fig:cross-traj-comparison-set})
These \emph{cross-trajectory co-observations} allow sub-teams to preserve the optimal unsecured trajectories (as shown in Figure \ref{fig:cross-traj-comparison-set}), as it does not require the entire \emph{sub-team} to maintain a close distance with other teams when co-observation occurs. To prevent potential situations where the entire sub-team is compromised and take the advantage of the decentralized blocklist protocol introduced in \cite{wardega2023byzantine} for additional Byzantine resilience, we always prefer \emph{cross-trajectory co-observation} when feasible.

@@

In \Cref{sec:ADMM-planning} we assumed that the number of robot is equal to the number of routes $n_p$.
In this section, we assume that we have a total of $n>n_p$ robots available. During initialization, $n$ robots are formed into \emph{sub-teams} by a time-varying partition $\cI(t)=\union_p \cI_p(t)$ of the robots where robots in each \emph{sub-team} $\cI_p$ share the same nominal trajectory. The planner introduced in \Cref{sec:ADMM-planning} is used to generate $n_p$ routes $\{q_p\}_{p=1}^{N_p}$ without reachability and co-observation constraints, where the state $q_p(t), p \in\{\cI_0,\ldots,\cI_{p}\}$ represents the reference position of the $i$-th sub-team at time $t\in\{0, \dots, T\}$. \zyang{should we mention admm is used here but other planners can be used as well?}

Expanding upon the concept of \emph{co-observation} and \emph{reachability}, \Cref{def:secured-plan} can be written as,
\begin{remark}\label{rmk:revised-security}
  a multi-robot trajectory is secured against plan-deviation attacks if there exist a co-observation plan such that the reachability region between each consecutive co-observation does not intersect with any forbidden regions.
\end{remark}

We capitalize on the presence of additional $n_p-n$ robots to enhance security.
In scenarios where at least one robot is mandated to adhere to the reference trajectory to fulfill essential tasks, these additional robots' initial focus are on performing co-observations within their respective sub-teams. Whenever necessary, they are designed to seamlessly transition between sub-teams. By deviating from their original sub-team's route and joining teams requiring heightened attention, they become instrumental in providing co-observations at critical junctures.

We refer to the new type of co-observations as \emph{cross-trajectory co-observations}. Our objective is to formulate a cross-trajectory co-observation strategy that fortifies an unsecured Multi-Robot System (MRS) trajectory, while minimizing the number of additional robots required.

To solve this problem, we model the planned trajectory $\{q_p\}_{p=1}^{N_p}$ as a directed \emph{checkpoint graph} $G_{q}=(V_{q}, E_{q})$. 
Vertices $V_{q}$ are a key locations in $\{q_p\}$ that requires attention.
The set $E_{q} = E_{t} \union E_{c}$ consists of directed edges connecting $V_i$ to $V_j$ if there exist possible paths between the corresponding locations.  Inspired by \cite{yu2013multi}, we show that additional robots in the sub-team can be formulated as flows in the checkpoint graph, transforming the co-observation planning problem into a network multi-flow problem that can be solved using general linear programming techniques as specialized solver.

@@

\subsection{Rapidly-exploring Random Trees}
In this section, we use the \rrtstar{} \cite{karaman2010incremental} algorithm to find paths from a single starting location to multiple destination points (i.e., the reference trajectories of other robots, in our method).

As a optimal path planning algorithm, \rrtstar{} returns the shortest paths between an initial location and points in the free configuration space, organized as a tree. We assume that the generated paths can be travelled in both directions (this is used later in our analysis). 
Key functions from \rrtstar{} that are also used during constructions of the checkpoint graph are
\begin{description}
\item[$\texttt{Cost}(v)$] This function assigns a non-negative cost (total travel distance in our application) to the unique path from the initial position to $v$ . 
\item[$\texttt{Parent}(v)$] This is a function that maps the vertex $v$ to $v'\in V$ such that $(v',v)\in E$.
\end{description}

 Notice that our objective is not to optimize specific constraints but to ascertain the existence of feasible paths. While the ADMM based solver \Cref{sec:ADMM-planning} presented earlier offers a broader range of constraint handling, the RRT* algorithm is better suited to the problem in this section. RRT*'s efficient exploration of the solution space, coupled with its ability to incorporate obstacle constraints, makes it a fitting choice for building the checkpoint graph.

@@

\subsubsection{Checkpoints}\label{sec:security-checkpoint}
Let $q_p\in\real{nT}$ be the MRS trajectory for sub-team $p$ with time horizon $T$, with $q_{p}(t) \in \real{n}$ be the reference waypoint at time $t$. 
\begin{definition}
Checkpoint $v_{pi}=(q_{pi},t_{i})$ is defined as a pair composed of a location $q_{pi}=q_{p}(t_i)$ and a time $t_{i}$ representing a co-observation required for sub-team $p$. 
\end{definition}
For convenience, let $\cI_{v_{i}}$ represent the sub-team that $v_{i}$ belongs to, \zyanghl{Is this necessary?}{and let $q_{v_{pi}} = q_{pi}$, $t_{v_{pi}}$ be the corresponding waypoint and time for checkpoint $v_{pi}$}.

@@

\begin{definition}
The set of $V_{p}=\{ v_{p0}, \dots ,v_{pT}\}$ (arranged in ascending order of $t_{v_{pi}}$) that can be used to secure the reference trajectory for sub-team $p$, if the reachability region between consecutive checkpoints $v_{pi}$ and $v_{p(i+1)}$ does not intersect with any of the forbidden areas. The requirement can be formally stated as $\mathcal{E}(q_{v_{pi}}, q_{v_{p(i+1)}}, t_{v_{pi}},t_{v_{p(i+1)}}) \intersect F = \emptyset$ for every $i$, where $F$ is the union of all forbidden areas as $F$. 
\end{definition}

We provide a heuristic approach to locate the checkpoints on given trajectories (an optimal solution would likely be NP-hard, while the approach below works well enough for our purpose). The approach is shown in \Cref{alg:checkpoint-gen}, and a toy example is shown in Figure \ref{fig:checkpoint-generate}. 
% We begin by adding the start and end locations $q_{p}(t_{0})$, $q_{p}(t_{e})$ to $V_{q}$, where $t_{0}=0$, and $t_{e}$ is the time horizon. Then, we search for the waypoint $q_{p}(t_{1})$ with the largest $t_{1} \in \{t_{0}, \dots ,t_{2}\}$ such that the reachability ellipsoid between $q_{p}(t_{0})$ and $q_{p}(t_{1})$ has no overlap with any of the forbidden areas, i.e. $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{1}), t_{0},t_{1})\intersect F = \emptyset$. Simultaneously, we search backward to find $q_{p}(t_{3})$ with the smallest $t_{3}\in \{t_{1}, \dots ,t_{2}\} $ that meets the reachability requirement $\mathcal{E}(q_{p}(t_{2}),q_{p}(t_{3}),t_{2},t_{3}) \intersect F = \emptyset$. $q_{p}(t_{1})$ and $q_{p}(t_{3})$ are then added to $V_{q}$. Afterwards, we set $t_{0}=t_{1}$, $t_{2}=t_{3}$ and repeat the same process. The search is stopped when $t_{0}>t_{2}$, or when $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{2}),t_{0},t_{2}) \intersect F = \emptyset$.

@@

\subsubsection{Cross-trajectory edges}\label{sec:cross-traj-edges}
After identifying checkpoints $\union_p V_p$ for all sub-teams, the next step is to search for cross-trajectory edges connecting all checkpoints to the respective trajectories. These cross-trajectory edges define viable paths between two reference trajectories, allowing robots to closely approach and collaborate with robots on a different reference trajectory for co-observations with the relevant sub-teams. Consequently, at least one endpoint of each edge must correspond to a security checkpoint $v_{p} \in \union_p V_p$. Assuming a single robot is transferred at a time between trajectories, the cross-trajectory edges must adhere to the reachability constraints detailed in ~\Cref{sec:reachability} to ensure that no deviations into forbidden areas can occur during trajectory switches.

>>>>

\rtron{Put proposition showing relation between ellipsoids and Definition 1 here.}

@@

To address the feasibility and performance trade-offs of the ADMM problem discussed in \Cref{sec:reachability-discussion}, we propose to form \emph{sub-teams} on each route, and setup additional co-observations both within the sub-team and across different sub-teams (as shown in \Cref{fig:cross-traj-comparison-set})
These \emph{cross-trajectory co-observations} allow sub-teams to obtain trajectories that are closer to the optimal (as shown in Figure \ref{fig:cross-traj-comparison-set}), because they do not require the entire \emph{sub-team} to meet with other teams for co-observations, thus providing more flexibility. Technically, \rtron{we can always create a trivial schedule that involves only co-observations between members of the same team; this, however, would make the solution more vulnerable in the case where multiple agents are compromised in the same team. While in this paper we explicitly consider only the single attacker scenario, multi-attacks can be potentially handled by taking advantage of the \emph{decentralized blocklist protocol} introduced in \cite{wardega2023byzantine}. For this reason, we setup the methods presented below to always prefer \emph{cross-trajectory co-observation} when feasible.}\rtron{Probably move this later.}

@@

In \Cref{sec:ADMM-planning} we assumed that the number of robot is equal to the number of routes $n_p$.
In this section, we assume that we have a total of $n>n_p$ robots available. During initialization, $n$ robots are divided into \emph{sub-teams}, i.e., we define a time-varying partition $\cI(t)=\union_p \cI_p(t)$ where robots in each sub-team $\cI_p$ share the same nominal trajectory. 
The planner introduced in \Cref{sec:ADMM-planning} is used to generate $n_p$ routes $\{q_p\}_{p=1}^{N_p}$ without reachability and co-observation constraints, where the state $q_p(t), p \in\{\cI_0,\ldots,\cI_{p}\}$ represents the reference position of the $i$-th sub-team at time $t\in\{0, \dots, T\}$. \zyang{should we mention admm is used here but other planners can be used as well? \rtron{Yes}}

Expanding upon the concept of \emph{co-observation} and \emph{reachability}, \Cref{def:secured-plan} can be written as, \rtron{Turn this into a proposition, and introduce it in Section II}
\begin{remark}\label{rmk:revised-security}
  a multi-robot trajectory is secured against plan-deviation attacks if there exist a co-observation plan such that the reachability region between each consecutive co-observation does not intersect with any forbidden regions.
\end{remark}

Since only $n_p$ robots are necessary to follow the initial routes, we capitalize on the remaining $n_p-n$ robots to enhance security. 
Each team will then have one robot to follow the reference trajectory and fulfill the basic sensing task, while the additional robots are dedicated to perform co-observations within their respective sub-teams. We assume the robots to be interchangeable by design, so that they can seamlessly transition between sub-teams.
\rtronhl{Not clear to me}{By deviating from their original sub-team's route and joining teams requiring heightened attention, they become instrumental in providing co-observations at critical junctures. }
We refer to this new type of co-observations as \emph{cross-trajectory co-observations}. Our objective is to formulate a cross-trajectory co-observation strategy that fortifies an unsecured MRS path finding solution, while minimizing the number of additional robots required.

To solve this problem, we model the planned trajectories $\{q_p\}_{p=1}^{N_p}$ with a directed \emph{checkpoint graph} $G_{q}=(V_{q}, E_{q})$. 
Vertices $V_{q}$ are a key locations in $\{q_p\}$ that requires \rtronhl{this term is not well defined}{attention}.
The set $E_{q} = E_{t} \union E_{c}$ consists of directed edges connecting $V_i$ to $V_j$ if there exist possible paths between the corresponding locations.  Inspired by \cite{yu2013multi}, we show that additional robots in the sub-team can be formulated as flows in the checkpoint graph, transforming the co-observation planning problem into a network multi-flow problem that can be solved using general \rtron{Mixed-Integer?}{Linear Programming} techniques.

@@

\subsection{Rapidly-exploring Random Trees}
In this section, we will use the \rrtstar{} \cite{karaman2010incremental} algorithm as a component of our method. In particular, \rrtstar{} will be used to find paths from a single starting location to multiple destination points, i.e., the reference trajectories of other robots, in our method.

As a optimal path planning algorithm, \rrtstar{} returns the shortest paths between an initial location and points in the free configuration space, organized as a tree. We assume that the generated paths can be travelled in both directions (this is used later in our analysis). 
While we refer the reader to \cite{karaman2010incremental,Choset:book05} for a full review of the \rrtstar{} algorithm, we highlight here two key functions that are also used during constructions of the checkpoint graph:
\begin{description}
\item[$\texttt{Cost}(v)$] This function assigns a non-negative cost (total travel distance in our application) to the unique path from the initial position to $v$ . 
\item[$\texttt{Parent}(v)$] This is a function that maps the vertex $v$ to $v'\in V$ such that $(v',v)\in E$.
\end{description}

 Notice that our objective is not to optimize specific constraints but to ascertain the existence of feasible paths. While the ADMM-based solver presented earlier in \Cref{sec:ADMM-planning} offers a broader range of constraint handling, \rrtstar's efficient exploration of the solution space, coupled with its ability to incorporate obstacle constraints, makes it a fitting choice for building the checkpoint graph.

@@

\subsubsection{Checkpoints}\label{sec:security-checkpoint}
Let $q_p\in\real{nT}$ be the MRS trajectory for sub-team $p$ with time horizon $T$, with $q_{p}(t) \in \real{n}$ be the reference waypoint at time $t$. 
\begin{definition}
Checkpoint $v_{pi}=(q_{pi},t_{i})$ is defined as a pair of a location $q_{pi}=q_{p}(t_i)$ and a time $t_{i}$ representing a co-observation required for sub-team $p$. 
\end{definition}
For convenience, let $\cI_{v_{i}}$ represent the sub-team that $v_{i}$ belongs to, \zyanghl{Is this necessary?}{and let $q_{v_{pi}} = q_{pi}$, $t_{v_{pi}}$ be the corresponding waypoint and time for checkpoint $v_{pi}$}.

@@

\begin{definition}\label{def:checkpoint vertices}
A set $V_{p}=\{ v_{p0}, \dots ,v_{pT}\}$ (arranged in ascending order of $t_{v_{pi}}$) is a valid set of vertices for the checkpoint graph if the reachability region between consecutive checkpoints $v_{pi}$ and $v_{p(i+1)}$ does not intersect with any of the forbidden areas.
\end{definition}
A sufficient condition for a set $V_p$ to satisfy \ref{def:checkpoint vertices} is that $\mathcal{E}(q_{v_{pi}}, q_{v_{p(i+1)}}, t_{v_{pi}},t_{v_{p(i+1)}}) \intersect F = \emptyset$ for every $i$, where $F$ is the union of all forbidden areas.

We provide a heuristic approach to locate valid checkpoint vertices on given trajectories (we do not consider an optimal selection, since finding this solution would likely be NP-hard, and the approach below works well enough for our purpose). The approach is shown in \Cref{alg:checkpoint-gen}, and a toy example is shown in Figure \ref{fig:checkpoint-generate}. 
% We begin by adding the start and end locations $q_{p}(t_{0})$, $q_{p}(t_{e})$ to $V_{q}$, where $t_{0}=0$, and $t_{e}$ is the time horizon. Then, we search for the waypoint $q_{p}(t_{1})$ with the largest $t_{1} \in \{t_{0}, \dots ,t_{2}\}$ such that the reachability ellipsoid between $q_{p}(t_{0})$ and $q_{p}(t_{1})$ has no overlap with any of the forbidden areas, i.e. $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{1}), t_{0},t_{1})\intersect F = \emptyset$. Simultaneously, we search backward to find $q_{p}(t_{3})$ with the smallest $t_{3}\in \{t_{1}, \dots ,t_{2}\} $ that meets the reachability requirement $\mathcal{E}(q_{p}(t_{2}),q_{p}(t_{3}),t_{2},t_{3}) \intersect F = \emptyset$. $q_{p}(t_{1})$ and $q_{p}(t_{3})$ are then added to $V_{q}$. Afterwards, we set $t_{0}=t_{1}$, $t_{2}=t_{3}$ and repeat the same process. The search is stopped when $t_{0}>t_{2}$, or when $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{2}),t_{0},t_{2}) \intersect F = \emptyset$.

@@

\subsubsection{Cross-trajectory edges}\label{sec:cross-traj-edges}
After identifying checkpoints $\union_p V_p$ for all sub-teams, the next step is to search for cross-trajectory edges connecting all checkpoints to the respective trajectories. 
\rtron{Add a definition mirroring Definition 8}
These cross-trajectory edges define viable paths between two reference trajectories, allowing robots to closely approach and collaborate with robots on a different reference trajectory for co-observations with the relevant sub-teams. Consequently, at least one endpoint of each edge must correspond to a security checkpoint $v_{p} \in \union_p V_p$. Assuming a single robot is transferred at a time between trajectories, the cross-trajectory edges must adhere to the reachability constraints detailed in ~\Cref{sec:reachability} to ensure that no deviations into forbidden areas can occur during trajectory switches.
